---
title: "Netflix Data Exploration: Filtering, Analysis, and Visualizations with R Studio by Andrew Romanof"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Data reference: https://www.kaggle.com/datasets/dgoenrique/netflix-movies-and-tv-shows?resource=download&select=titles.csv

IMDB (Internet Movie Database) and TMDB (The Movie Database) are two seperate unrelated databases that store movies and tv shows information such as title, release date, actors, directors, rating, etc. I want to compare the average rating of both databases for tv shows and films to see which one people enjoy more and to see if there is a difference in rating between the two databases.

The results below show that people like tv shows more than movies on average and that both databases share the same score on average (slightly larger score on TMDB). Since both databases are unrelated, I made a graph for each database.

```{r}
# load in library
library(tidyverse)
# load in data
data <- read.csv("titles.csv")
# create dataframe for imdb
imdb <- data %>% group_by(type) %>% summarise(mean_imdb = mean(imdb_score, na.rm = TRUE))
# create dataframe for tmdb
tmdb <- data %>% group_by(type) %>% summarise(mean_tmdb = mean(tmdb_score, na.rm = TRUE))
# plot imdb
ggplot(imdb, aes(x = type, y = mean_imdb, fill = type)) +
  geom_bar(stat="identity", colour = "#33FFFF") +
    labs(x = "Type", y = "Average IMDB Rating", title = "Average IMDB Rating for shows and movies on Netflix",
       fill = "Type") +
  theme(plot.title = element_text(face = "bold.italic", size = 16, hjust = 0.5), 
        panel.grid.major = element_line(colour = "grey", linewidth = 1), 
        panel.background = element_rect(fill = "beige", colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "beige", colour = "black"), 
        ) +
  scale_y_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)) + 
  scale_fill_manual(values = c("coral", "deepskyblue"))
  
# plot tmdb
ggplot(tmdb, aes(x = type, y = mean_tmdb, fill = type)) +
  geom_bar(stat="identity") +
    labs(x = "Type", y = "Average TMDB Rating", title = "Average TMDB Rating for shows and movies on Netflix",
       fill = "Type") +
  theme(plot.title = element_text(face = "bold.italic", size = 16, hjust = 0.5), 
        panel.grid.major = element_line(colour = "grey", linewidth = 1), 
        panel.background = element_rect(fill = "beige", colour = "black", linewidth = 1),
        legend.background = element_rect(fill = "beige", colour = "black"), 
        ) +
  scale_y_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)) + 
  scale_fill_manual(values = c("coral", "deepskyblue"))
```

Since the previous figure establishes that on average, IMDB AND TMDB scores are the same, this next figure will be using only IMDB score.

I want to see if there is a relation between the amount of seasons a TV show has and its IMDB rating. To do this, I'm goingt to break the amount of seasons to different categories: Short, Medium, Large where short means that the tv show has 1-3 seasons, medium means that the tv show has 4-8 season and that larrge means the tv show has 9+ seasons.

The below figure shows that on average, IMDB rating goes up as the amount of seasons go up (which is opposite to what I expected to find as I thought shows would get worse as more seasons get added)


```{r}
# filter to have the dataframe contain only shows
shows <- filter(data, str_detect(data$type, "SHOW"))
# make a new category to classify shows as short, medium or large
shows <- mutate(shows, season_size = if_else(shows$seasons <= 3, "Short", 
                                           if_else(shows$season > 3 & shows$seasons < 9, "Medium", "Large")))
# group the shows by season size and find the average imdb score of them
shows <- shows %>% group_by(season_size) %>% summarise(mean_imdb_score = mean(imdb_score, na.rm = TRUE))
# plot the results
ggplot(shows, aes(x = mean_imdb_score, y = season_size, fill = season_size)) +
  geom_bar(stat="identity", show.legend = FALSE) +
    labs(y = "Amount of seasons", x = "Average IMDB Rating", title = "Average IMDB Rating for TV Shows by season size on Netflix",
       fill = "Type") +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)) +
  scale_y_discrete(labels = c('Short' = "Short \n (3 or less)", 'Medium' = "Medium \n (Between 4 and 8)",
                                'Large' = "Large \n (9 or more)")) +
  theme(axis.ticks.length=unit(0.3, "cm"), axis.text = element_text(size = 10, face = "bold"),
        panel.background = element_rect(colour = "black", linewidth = 1.5),
        plot.background = element_rect(colour = "black", linewidth = 2, fill = "beige"),
        plot.title = element_text(hjust = 1, face = "bold.italic"))
```
Now lets see how the IMDB score does based on countries...

The figure below shows the average IMDB rating for each country on a world map. Note: not every country is listed and some data points were missing.

We can conclude that the highest average rated movies/tv shows come from China, Tanzania (Country in Africa), Pakistan. Although, production rate in Tanzania might not be as high compared to USA, it seems that a movie from there will be more likely to be higher rated than a movie from USA showing that quality over quantity is better. It also shows that average IMDB ratings are relatively close to each other (lowest being just under 5 and the highest being just above 7) and from previous figures, the average IMDB rating overall is around 7 so the results also imply that if a country has a low average IMDB rating, then it probably hasn't produced much.


```{r}
# import library countrycode to convert country abbreviations to names
library(countrycode)

# edit the data a bit to remove quotations and brackets
data$production_countries <- str_replace_all(data$production_countries, "'", "")
data$production_countries <- str_remove_all(data$production_countries, fixed("["))
data$production_countries <- str_remove_all(data$production_countries, fixed("]"))

# remove warning message and convert the country abbreviations to names in a new column
suppressWarnings(data$country <- countrycode(data$production_countries, origin = "iso2c", destination = "country.name"))
# replace "United States" to "USA" so that it can be mapped
data$country <- str_replace_all(data$country, "United States", "USA")

# find average imdb score by country
countrydata <- data %>% group_by(country) %>% summarise(mean_imdb = mean(imdb_score, na.rm = TRUE))

# plot on world map
world_map <- map_data("world")
ggplot(countrydata) +
    geom_map(data = world_map, map = world_map,
    aes(map_id = region), fill = "white", colour = "black") +
    geom_map(map = world_map,
    aes(map_id = country, fill = mean_imdb), colour = "black") +
    expand_limits(x = world_map$long, y = world_map$lat) +
    labs(x = "Longitude", y = "Latitude", title = "Average IMDB Score by Country",
       subtitle = paste0("(", length(unique(countrydata$country)), " countries)"), fill = "Average IMDB Score")
```


